---
title: 'Task 16: The US Grocer (KR)'
author: "McKay Davis"
output: 
  html_document:
    code_folding: hide
    theme: united
    keep_md: TRUE
---

```{r setup, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE)

library(tidyverse)
library(dygraphs)
library(tidyquant)
library(xts)
library(DT)

KR <- tq_get("KR", get = "stock.prices", from = "2012-01-01") %>%
  xts(order.by = .$date)
```

##Kroger (KR) Stock Prices

```{r, fig.align = 'center', fig.width = 12, fig.height = 6}
datatable(KR, colnames = c("Time" = 1), extensions = 'Scroller', options = list(
  deferRender = TRUE,
  scrollY = 200,
  scroller = TRUE
)) %>% 
  formatStyle('adjusted',  color = 'darkblue', backgroundColor = 'yellow', fontWeight = 'bold')

```

<br/>

```{r, fig.align = 'center', fig.width = 9, fig.height = 6}
KR <- KR[,7]
dygraph(KR, main = "Kroger Adjusted Stock Prices") %>% 
  dySeries("adjusted", label = "Adjusted") %>% 
  dyAxis("x", drawGrid = FALSE) %>%
  dyAxis("y", label = "Adjusted") %>%
  dyOptions(axisLineColor = "navy", 
            gridLineColor = "lightblue",
            fillGraph = TRUE) %>% 
  dyAnnotation("2014-2-6", text = "A", tooltip = "Unemployment Benefit Cashed, Less than Anticipated\nLeads to increase") %>%
  dyAnnotation("2016-11-9", text = "B", tooltip = "Trump Wins Election") %>% 
  dyAnnotation("2017-6-14", text = "C", tooltip = "Kroger Pricing a Problem as Walmart Competes") %>% 
  dyRangeSelector(height = 20)

```

<br/>

The plot above shows the adjusted KR stock prices for January 2012 - date. A couple points are labeled showing events of interest and their effect on the stock prices.

######Details of Plot:
I chose to get rid of the grid in favor of light horizontal lines as the main interest in the plot is the adjusted stock price. I filled the graphic for asthetic purposes. The slide bar is added for convenience. I did not label the x-axis because it is obvious what it is.

####More on Annotations:
<p>A: [Best Gains of The Year](http://money.cnn.com/2014/02/06/investing/stocks-markets/index.html)</p>
<p>B: Donald Trump wins the presidential election which leads to overall increase.</p>
<p>C: [Kroger Market Fails](https://www.forbes.com/sites/maggiemcgrath/2017/06/15/kroger-sinks-16-as-first-quarter-sales-fall-and-forward-guidance-is-slashed/#37259468726d)</p>



